---
title: "bericht pdf"
output: pdf_document
date: "2023-07-15"
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The advancement in the creation of new drugs to combat chronic diseases and pathologies that today represent a global priority in health research, is one of the great challenges of medicine and biotechnology nowadays.
This is especially evident in diseases such as cancer, due to its high incidence and mortality rate (Carcas, 2014).
A large number of potential antitumor drugs tend to fail in clinical trial stages, as they are not sufficiently effective or safe for human use, and simply do not reach the stage of being approved for distribution on the market.
Moreover, the approval processes are remarkably long, lasting between 10 and 17 years.
(Ioakeim-Skoufa et al., 2023).
Likewise, the most effective and novel treatments tend to be expensive, making them inaccessible for oncological patients from middle and low-income countries.
This situation has generated a growing interest in the search for more affordable, less toxic, and faster-to-obtain alternatives.
An innovative strategy is "drug repurposing", which consists of re-evaluating existing and approved drugs to treat diseases different from cancer, which could have an antitumor effect.
Once their potential against cancer is discovered, clinical trials can be initiated in patients, reducing both the time and final cost to the public, as these drugs are already available in the market (Ioakeim-Skoufa et al., 2023) Cancer is among the deadliest illnesses, characterized by a remarkably high death rate Currently, one of the types of cancer that persists as a serious health problem is gastric cancer, despite a downward trajectory in global incidence.
It is recognized as the second leading etiology of cancer-related mortality globally and registers as the fifth most prevalent cancer diagnosis (Carcas, 2014).
Among the different types of gastric cancer, adenocarcinoma is the most common (Cebrián, Concha, & Fernández-Urién, 2016).
In the context of early-stage tumors, the first therapeutic options are endoscopic resection, surgery, and or an integrated regimen of chemotherapy or chemoradiation.
However, the treatment of advanced or metastatic gastric cancer is not very far reaching, and normally in this stage the overall survival time rate in this group is less than one year (Carcas, 2014).
This landscape highlights the urgency to develop and explore new therapeutic alternatives, such as drug repurposing, to improve the survival and quality of life of patients with gastric cancer.

In this study, our main objective is to systematically explore the responses of a broad spectrum of drugs in a diverse range of cancer cell lines, with special emphasis on gastric cancer.
With the help of additional information on factors such as gene expression, genetic alterations and target molecules, we aim to identify potential therapeutic agents for the treatment of cancer through the analysis of intermolecular relationships.
More specifically, we aim to identify distinct subgroups of these potential treatments based on their molecular properties.
Ultimately, this research aims to contribute to the evolving field of drug repositioning as a viable strategy to advance gastric cancer therapeutics.

# Material and Methods

### 2.1 Data Sets

For this project, we were provided with two data sets, consisting of seven data frames in total.
We can divide them into three different categories.
The first data frame, called "prism", provided essential information regarding the response of 481 different cell lines to various treatments in terms of cell growth.
The second data frame, "prism.treat," contained details about the treatments themselves, such as the drugs used, associated pathways and target molecules.
The third data frame, "prism.cl", provided us with information about the 481 cell lines.
Additionally, we utilized three other data frames to examine gene expression and alterations in the cell lines "prism.exp" contained data on the levels of gene expression for 19,177 different genes in the cancer cell lines.
The "prism.cnv" data frame provided gene copy values, indicating gene amplifications or deletions.
The "prism.snv" data frame contained information about observed mutation types in the samples.
Lastly, the "prism.achilles" data frame included gene knockdown scores, indicating the impact of reducing gene activity on cell growth.

## 2.2 Data cleanup

Remove the low variance

Gastric cancer cell lines extraction

Remove the NA values

## 2.3 Identifying the effects of treatments on cell growth of gastric cancer cell lines + breast cancer

We conducted a comprehensive analysis of drug responses in gastric cancer cell lines, with a focus on reducing high-dimensional data.
Therefore, we applied  Principle Component Analysis (PCA) to reduce the dimensionality of the dataset prism.
Furthermore, we performed Pearson correlation on each individual principal component to assess statistically the strength and the direction of the linear relationship between the drugs and the cell growth of gastric cancer cell lines.
Combining PCA and Pearson correlation provided us information about the associations between drug treatments and their effects on cell growth, helping us identify potential targets and molecular pathways(Goncalves et. al.).
We also conducted Pearson correlation to examine the relationship between drug concentration and drug response. 

## 2.4 Finding clusters of the cell lines based on gene expression and drug response

Principal component analysis was conducted, using data from prism and prism.exp.
By calculating the variance of each row in the transposed data frame of prism and prism.exp, the variance of the drug response or the gene expression in each analyzed cell line was obtained.
Plotting the values of variance of all drugs and genes in a histogram led to a better overview of the variability of the drug response and the expression of genes.
Furthermore, did the visualization of the variances help to select a threshold of minimum variance for the gene expression or drug response in cell lines.
Only drugs and genes of high variance were chosen to perform principal component analysis to reduce the data only for drugs and genes which explain the genetic differences of the cell lines.
To visualize the results obtained through the use of PCA, two plots were created: a scree plot and a scatter plot, respectively for prism and prism.exp.
The scree plot demonstrates the variance explained by each principal component, providing an insight into the contribution of each component towards the overall variation in the data.
On the other hand, the scatter plot displays the scores of PC1 and PC2 for each cell line.
This plot allows for the examination of the relationships and patterns among the cell lines based on their scores on these two principal components.
Color coding each cell line corresponding to the cancer type displays how cell lines of the same cancer type cluster together by analyzing for example the genes that show the highest variance of expression in the cell lines. 

Since the scatter plot of the PCA only indicates possible clusters of the cell lines, K-means clustering was applied to group the cell lines together based on the PCA results.
To find the optimal amount of clusters, the elbow and silhouette method was used.
Plotting the sum of squared distances between data points and their assigned cluster centers (also known as the within-cluster sum of squares or WCSS) for different numbers of clusters lead to the plotXX in which the so called "elbow" which indicates the optimal amount of clusters, is obtained at k=3. 
The silhouette method takes into account the average distance to other points within the same cluster and the average distance to points in the nearest neighboring cluster.
Resulting from analyzing the silhouette coefficient for possible values of k, the optimal amount of clusters seem to be either three or five when analyzing the gene expression for example XX.

## 2.5 Fold Change Analysis: Evaluating Non-Oncology Versus Oncology Treatment Responses

A fold change analysis was conducted to compare the responses of potential non-oncology treatments to oncology treatments as the control group.
Our focus was on treatments relevant to gastric or gastrointestinal indications, or in other analysis, oncology in general so all treatments matching these criteria were identified from the dataset PRISM.treat.
A similar approach was used to identify the corresponding non-oncology treatments.
In order to ensure a more reliable analysis, data cleaning and processing were a crucial component of our workflow.
Missing values were replaced by the mean response for each treatment.
The average response for each cell line for the various oncology and non-oncology therapy groups was then determined independently.
So consequently, a fold change was then computed for each cell line, defined as the ratio of the average non-oncology treatment response to the average oncology treatment response.
This provided a measure of the relative response to non-oncology versus oncology treatments for each cell line.
A pseudo count was added to both the numerator and the denominator during the fold change calculation to avoid division by zero and stabilize the ratio, especially for small values.

## 2.6 Finding drug-gene pairs by calculating the correlation of the drug response and its target gene expression or gene copy number

Obtaining the target of each drug of which the drug response is shown in the prism data frame by using the prism.treat data frame and analyzing the correlation between the drug response and the expression as well as the gene copy number of the drugs target in the cell line.
The drugs that were identified through PCA as contributing the most to the reduction of cell growth in gastric cancer cell lines were utilized for this analysis.
Lists were created that contain all calculated correlations of the drug responses of a drug in gastric cancer cell lines and the gene expression, evaluated by the TPM values shown in prism.exp data frame, or the gene copy numbers taken from the prism.cnv data frame  of the drugs targets in the cell lines.
Then the drug-gene pairs were selected showing the highest absolute values as for their correlation of response and expression or gene copy number values.
These were plotted in a scatter plot to visualize the linear correlation between them.

# **Results**

## 3.1 Data reduction based on drug response of gastric cancer cell lines

First of all, our primary objective was to reduce the dimensionality of the prism dataframe, focusing specifically on gastric cancer cell lines by applying Principal Component analysis(PCA).
Through the analysis, we performed 2 different PCAs and generated  scatter plots to visualize the results.
In the first scatter plot, each dot represents a gastric cancer cell line (Supplementary Materials), while in the second scatter plot each dot represents drug( Figure ).

```{r echo=FALSE}
load("~/Uni Heidelberg Molekulare Biotechnologie/4.Semster/Bioinfo/cellline_datasets.rda")
load("~/Uni Heidelberg Molekulare Biotechnologie/4.Semster/Bioinfo/prism_datasets.rda")
```

```{r echo=FALSE,fig.width = 4, fig.height = 4}

gastric.cancer.celllines <- subset(prism.cl, lineage == "gastric")


gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID



prism_gastric <- prism[rownames(prism) %in% gastric.cancer.celllines.v, ]


prism_gastric_cleaned_treatments <- prism_gastric[, colSums(is.na(prism_gastric)) == 0]
prism_gastric_cleaned_treatments_t<- t(prism_gastric_cleaned_treatments)

pca_result_gastric_cleaned_treatments_t<- prcomp(prism_gastric_cleaned_treatments_t)
pca_values_gastric_cleaned_treatments_t <- pca_result_gastric_cleaned_treatments_t$x # Principal component scores (transformed data)
pca_variances_gastric_cleaned_treatments_t <- pca_result_gastric_cleaned_treatments_t$sdev^2  # Variance explained by each principal component
pca_proportions_gastric_cleaned_treatment_t <- pca_variances_gastric_cleaned_treatments_t / sum(pca_variances_gastric_cleaned_treatments_t)  # Proportion of variance explained by each principal component
pca_loadings_gastric_cleaned_treatments_t <- pca_result_gastric_cleaned_treatments_t$rotation  # Loadings (coefficients) of each variable on each principal component

plot(pca_values_gastric_cleaned_treatments_t[, 1], pca_values_gastric_cleaned_treatments_t[, 2], xlab = "PC1", ylab = "PC2", main = "PCA scatter plot for drug response of
     gastric cancer cell lines")


```

Subsequently, we selected six from seventeen principle components, based on their contribution to the overall variance being more than 10 percent.
Then, we computed the Pearson correlation of six principal components to the gastric cancer cell lines.
This Pearson correlation analysis visualizes the contribution of each principal component to the cell growth of gastric cancer cell lines.

```{r echo=FALSE, fig.width = 5, fig.height = 2}

#Create a data frame with the correlation values
library(ggplot2)
correlation_data <- data.frame(CellLine = 1:17,
                               PC1 = c(-0.922615177, -0.950194073,   -0.927977386, -0.94910213, -0.9152670, -0.854801869, -0.95349076, -0.92154645, -0.92013903, -0.77537470,-0.950086951, -0.93543329, -0.93709137, -0.9250477213, -0.950361834,-0.93871668, -0.94822669),  
                               PC2 = c(0.083180285, -0.095496631,  0.179019069,   0.03838714, -0.1386250, 0.345058466, -0.11534765,  0.20474041, -0.06588813,  0.43218037, -0.088239721, -0.14684554, -0.00350409,  0.1180914557, -0.065329501,-0.02260016, -0.06579373), 
                               PC3= c(-0.063771576,  0.154186354, -0.007993429,  0.11713865, -0.2703927,-0.152188017, -0.07227995,  0.04222941,  0.04007063,  0.16919529, 0.145417339,  0.09744200, -0.06908235, -0.0006997035, -0.025131615,-0.04229883, -0.04697155),
                              PC4= c(0.001032055, -0.030713988,  0.157945987,  0.02180625,  0.1161595,-0.319057038, -0.02522252,  0.06795268, -0.12543977,  0.14860319,-0.034134148, -0.03687671,  0.10518056,  0.1658635330,  0.007893927,-0.04026534, -0.04409653),
                              PC5= c(-0.133581302,  0.001271816,  0.036998181, -0.04956825, -0.1377278, 0.001838434,  0.04777600, -0.05017946, -0.21094701, -0.05346703,0.001838434,  0.04777600, -0.05017946, -0.21094701, -0.05346703,0.20053979,  0.17713578),
                              PC6= c(0.001307767, -0.030989500, -0.050914709, -0.04882622,  0.0351149,-0.103958062,  0.04898928, -0.04865855,  0.21335588,  0.19243498,0.017044198, -0.13232765, -0.12824958,  0.0502075589, -0.159971620, 0.09320630,  0.11132276))
                              
                               
# Convert the data from wide to long format
correlation_data_long <- tidyr::gather(correlation_data, PC, Correlation, -CellLine)
my_theme <- theme_minimal() +
  theme(plot.title = element_text(size = rel(0.6)),  # Adjust the size of the plot title
        axis.title = element_text(size = rel(0.4)))  # Adjust the size of the axis labels

# Create the bar plot using ggplot
ggplot(correlation_data_long, aes(x = PC, y = Correlation, fill = factor(CellLine))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Pearson Correlation of Principal Components 
       of Gastric Cancer Cell Lines",
       x = "Principal Components",
       y = "Pearson Correlation") +
  scale_fill_discrete(name = "Gastric Cancer Cell Lines") +
  ylim(-1, 1) +
  theme_minimal() +
  theme(legend.text = element_text(size = 6),  # Adjust the size of legend text
        legend.key.size = unit(0.4, "cm"))



```

All the drugs contributed positively to PC1, which contributed 86.34% to the overall variance.
These drugs demonstrated a consistent negative effect on cell growth of all gastric cancer cell lines.
Specifically, there were 678 drugs identified as positively contributing  to the PC1.
Therefore, we selected a threshold as 20 for PC values and we continued the further analysis with 66 selected drugs, which contributed highly to PC1.

## Identifying top targets and mechanism of action for drugs inducing strong negative cell growth in gastric cancer cell lines

With the aim of identifying the target molecules and the mechanism of actions we reduced the prism.treat dataframe by extracting these 66 selected drugs with different concentrations.
To focus our analysis, we set 2 as threshold for the target and moa frequency.
Moreover, we plotted 2 histograms which visualizes the top targets and mechanism of actions associated with the selected drugs.

```{r include=FALSE}
pc1_selected_drugs <- rownames(pca_values_gastric_cleaned_treatments_t)[pca_values_gastric_cleaned_treatments_t[, "PC1"] > 20]
prism_treat_new <- prism.treat[rownames(prism.treat) %in% pc1_selected_drugs, ]
prism_treat_new2 <- subset(prism_treat_new, dose == "10")
big_string <- paste(prism_treat_new2$target, collapse = ",")
target_vector <- unlist(strsplit(big_string, ","))
target_vector <- gsub(" ", "", target_vector)
target_frequency <- as.data.frame(table(target_vector))
target_freq_dataframe_selected <- subset(target_frequency, Freq > 2)
# Remove the row with "NA" in target_vector column
target_freq_dataframe_selected <- target_freq_dataframe_selected[target_freq_dataframe_selected$target_vector != "NA", ]
# Create a bar plot using ggplot2
bar_plot_target <- ggplot(target_freq_dataframe_selected, aes(x = reorder(target_vector, -Freq), y = Freq, fill = target_vector)) +
  geom_bar(stat = "identity") +
  labs(title= " Target Molecules of Selected Drugs ", x = "Target", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylim(0, 15)  
bar_plot_target <- bar_plot_target + scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))
print(bar_plot_target)
```


```{r include=FALSE}
# Seperate the strings of the column target
big_string <- paste(prism_treat_new2$moa, collapse = ",")
moa_vector <- unlist(strsplit(big_string, ","))
moa_vector <- gsub(" ", "", moa_vector)
moa_frequency <- as.data.frame(table(moa_vector))
# create a new dataframe of target frequency and select those which have more than 1 freq:
moa_freq_dataframe_selected <- subset(moa_frequency, Freq > 2)

# Create a bar plot of moa
bar_plot_moa <- ggplot(moa_freq_dataframe_selected, aes(x = reorder(moa_vector, -Freq), y = Freq, fill = moa_vector)) +
  geom_bar(stat = "identity") +
  labs(title= "Mechanism of Actions 
       of Selected Drugs", x = "moa", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylim(0, 15)  # Adjust the range as needed

##Adjust the x-axis labels
bar_plot_moa <- bar_plot_moa + scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))


print(bar_plot_moa)
```

```{r echo=FALSE, fig.width = 8, fig.height = 4}
library(ggplot2)
library(cowplot)

# Create the first bar plot
bar_plot_target <- ggplot(target_freq_dataframe_selected, aes(x = reorder(target_vector, -Freq), y = Freq, fill = target_vector)) +
  geom_bar(stat = "identity") +
  labs(title= "Target Molecules of Selected Drugs", x = "Target", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 15) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))

# Create the second bar plot
bar_plot_moa <- ggplot(moa_freq_dataframe_selected, aes(x = reorder(moa_vector, -Freq), y = Freq, fill = moa_vector)) +
  geom_bar(stat = "identity") +
  labs(title= "Mechanism of Actions of Selected Drugs", x = "moa", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 15) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))

# Arrange the plots side by side
plot_grid(bar_plot_target, bar_plot_moa, ncol = 2)


```

The first histogram revealed that  HDAC (Histone Deacetylase) emerged as the most commonly targeted molecule among the 66 selected drugs.
Consequently, HDAC inhibition was identified as the dominant mechanism of action among the selected drugs. 

## Identifying missense mutation in HDAC as an important factor for cell growth and proliferation in gastric cancer

Furthermore, we extracted the genes associated with the top targets of selected drugs namely *HDAC1, HDAC2, HDAC3, HDAC4, HDAC5, HDAC6, HDAC7, HDAC8* and *HDAC9* from prism.snv dataframe.
Moreover, we plotted a histogram, which visualized the distribution of the mutation type frequencies in these selected genes (Figure ).
The distribution of mutation types indicates that missense and silent mutation in HDAC genes contribute to the dysfunctional Histone deacetylase activity, which contributes to cell growth and proliferation in gastric cancer.
Further discussion on these analysis will be presented in the discussion part.

```{r echo=FALSE, fig.width = 4, fig.height = 3}
selected_genes <- c("HDAC1", "HDAC2", "HDAC3", "HDAC4", "HDAC5", "HDAC6", "HDAC7", "HDAC8", "HDAC9")

selected_prism_snv <- prism.snv[prism.snv$Hugo_Symbol %in% selected_genes, ]
# Count the frequency of each mutation type
mutation_counts <- table(selected_prism_snv$Variant_Classification)

# Create a data frame with mutation types and their counts
mutation_data <- data.frame(Mutation_Type = names(mutation_counts),
                            Count = as.numeric(mutation_counts))

# Sort the data frame by mutation type counts in descending order
mutation_data <- mutation_data[order(mutation_data$Count, decreasing = TRUE), ]

# Plot the distribution of mutation types
ggplot(mutation_data, aes(x = reorder(Mutation_Type, Count), y = Count)) +
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(title = "Distribution of Mutation Types 
       of Selected Genes",
       x = "Mutation Type",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### **Exploring the Correlation Between HDAC Inhibitor Concentration and Gastric Cancer Cell Growth**

Through our investigation, we identified six HDAC inhibitors that exhibited strong inhibitory effects on cell growth.
To gain deeper insights into the relationship between drug concentration and drug response, we conducted a comprehensive analysis by calculating the correlation between these two factors.
To facilitate this analysis, we extracted the relevant information for these six treatments, including their respective concentrations and their corresponding effects on cell growth in gastric cancer cell lines.
By creating a new data frame specifically for these treatments, we were able to perform correlation calculations for each cell line separately.
To visualize the results, we created a heat map, which effectively captured the relationship between drug concentration and cell growth response(Figure ).
The heat map highlighted a robust negative correlation between the concentration of the HDAC inhibitors and cell growth.
In other words, as the dosage of the inhibitors increased, the drug influence on the cell growth became stronger, leading to a more pronounced negative impact.

## **Cell lines of different cancer types form different clusters based on their drug response and gene expression**

To analyze the similarities and differences regarding drug response and gene expression among cell lines, principal component analysis (PCA) and K-means clustering were performed.
The log fold change values from the "prism" dataset and TPM values from the "prism.exp" dataset were used, which contained data from cell lines representing five different cancer types: gastric, breast, bone, liver, and kidney.

The PCA scatter plot of drug response is showing similar drug response patterns among all of the examined cancer subtypes.
Most of the cell lines formed one cohesive cluster, which can be further divided into smaller clusters by applying K-Means clustering.
Notably, these similarities in drug response emerge, even when analyzing drugs with high variance.
In other words, these drugs exhibit similar influence on cell growth irrespective of the particular cancer subtype.
This suggests that these drugs may have common mechanisms of action and common targeted pathways. 

On the other side, the PCA scatter plot of gene expression revealed that cell lines of the same cancer type tend to cluster together, indicating similarities in their gene expression profiles.
However, there were noticeable differences observed within certain cancer types, such as liver and bone cancer cell lines, resulting in a broader distribution in the scatter plot.
Furthermore gastric and breast cancer cluster extremely together, forming their own clusters even when forming different amounts of clusters which indicates a strong similarity of gene expression that may suggest similar drug responses as well. 

## **Shared Targets and Mechanisms of Action in Gastric and Breast Cancer: Comparative Analysis**

In order to validate the observed clustering between gene expression patterns of gastric cancer and breast cancer and their similar drug response, we conducted further analysis to substantiate the significance of this finding.
To achieve this, we performed principal component analysis (PCA) exclusively on the prism data set, focusing only on breast cancer cell lines, with the objective of identifying key targets and potential mechanisms of action and to compare them with gastric cancer.
The methods applied in this analysis was consistent with the approach explained in the initial sections of the report (see first and second results).
Subsequently, we generated two histograms.
The first histogram presented the targets that exhibited a frequency of occurrence greater than two, while the second histogram visualized the most commonly observed mechanisms of action.
It is important to note that this analysis focused exclusively on drugs that made the most substantial contributions to PC1.
Remarkably, our analysis revealed the presence of three specific and noteworthy targets that are shared between gastric cancer and breast cancer subtypes: EGFR, HDAC1, and mTOR.
Additionally, upon examining the second histogram, we observed a convergence of mechanisms of action.
Specifically, HDAC inhibitors and FLT3 inhibitors emerged as the top three mechanisms common to both cancer subtypes.
These findings demonstrate that breast and gastric cancer exhibit remarkable similarities in terms of shared targets, as well as common mechanisms of action.
The importance of this result will be further observed in the discussion part. 

## **Analyzing the correlation between the response of drugs and the CNVs as well as the gene expression of the drugs target in gastric cancer cells**

Correlations between drug response and gene expression or CNVs of drug targets in gastric cancer cells identified two drug-gene pairs for further analysis: BRD-K56957086-001-06-3/HDAC1 and BRD-K35329391-334-01-7/BCR.
These pairs demonstrated strong correlations between the drug response and both the gene expression and CNVs of their respective targets.

As before mentioned HDAC inhibitors seem to be a promising group of repurposing drugs for gastric cancer as the graphs (Figures) depicting the drug response of BRD-K56957086-001-06-3 against HDAC1 gene expression or CNVs in gastric cancer cell lines confirm.
The reduction of cell growth is dose-dependent of BRD-K56957086-001-06-3 since the dots of lower doses are clustered around zero, meaning no effect, while increasing the dose lead to a reduction of cell growth in all gastric cancer cell lines (Figures).
Same can be observed when looking at the scatter plots (Figures) visualizing the relationship between the drug response of  BRD-K35329391-334-01-7 and the gene expression or the CNVs of its target BCR.
Compared to HDAC inhibitor a much higher concentration is needed so the effect of BRD-K35329391-334-01-7  is guaranteed since in some cell lines the second highest dose of 2.5 mM has nearly no effect on cell growth, as the data points are located around zero.
At a concentration of 10 mM, a strong reducing effect on cell growth can be observed in all gastric cancer cell lines.

The correlations between drug response of BRD-K56957086-001-06-3 and HDAC1 gene expression vary when increasing the drug dosage (Figure).
In fact, there is no consistent pattern or trend in the correlations as the dose of BRD-K56957086-001-06-3 increases.
Contrarily, as the drug dose of BRD-K56957086-001-06-3 increases, the correlations with HDAC1 CNVs consistently increase (Figure).
This finding supports the hypothesis that the drug's effect in gastric cancer cells is indeed linked to the CNVs of its target gene, HDAC1.
For very low dosages, the overall effect on the cell lines is minimal, likely due to the fact that the effective dose of the drug is not reached, thereby hindering a clear expression of the correlation between drug response and HDAC1 CNVs.
Similar observations can be made when examining the change of the correlation between the drug response of BRD-K35329391-334-01-7 and the gene expression or the CNVs of its target BCR when increasing the dosage of the drug(Figures).
However the drug response of BRD-K35329391-334-01-7 and the CNVs of BCR in gastric cancer cells show a negative correlation for all dosages of the drug (Figure).
Lower CNVs suggest a weaker reduction in cell growth, while higher CNVs indicate an increased effectiveness of the drug in reducing cell growth in gastric cancer cells. 

## **Creating prediction models for the drug response of two specific drugs based on the CNVs of their target gene**

Since the correlations of both drugs with their target CNVs were higher compared to the correlation between drug response and target gene expression and also showed a constant increase when increasing the dosage, linear regression was applied to further analyze the relationship between the drug response of BRD-K56957086-001-06-3 and BRD-K35329391-334-01-7 and the CNVs of HDAC1 and BCR in gastric cancer cells.
The objective is to develop a prediction model for estimating the drug response of a specific drug, using the copy number variations (CNVs) of the drug's target gene as the independent variable and perform linear regression.
The model was created only for the highest dosage of the drugs namely 10 mM.
XX

The equation which describes the created prediction model of BRD-K56957086-001-06-3 (Y-dependent variable) and the CNVs of HDAC1 (X- independent variable) in gastric cancer cells states as follows: Y = -10.285 + 4.347X.
The coefficient is statistically significant (p-value = 0.0301), suggesting a positive relationship between the drug response of BRD-K56957086-001-06-3 and CNVs of HDAC1.

The intercept is estimated to be -10.285 with a standard error of 1.732.
It is also statistically significant (p-value \< 0.0001).
The model has a multiple R-squared of 0.2766, indicating that approximately 27.66% of the variance of the drug response is explained by the model.
The F-statistic (5.736) is significant, indicating that the model as a whole is statistically significant.
The residual standard error is 0.8758, which represents the average deviation of the observed values from the predicted values.

The equation explaining the results of linear regression analysis with BRD-K35329391-334-01-7(Y-dependent variable) and the CNVs of BCR (X- independent variable) in gastric cancer cells is:Y = -4.2900 - 3.6193 X.
Both the intercept and the coefficient are statistically significant, indicated by the corresponding p-values (\< 0.05).
The model explains approximately 46.77% of the variance of the drug response, as indicated by the multiple R-squared value of 0.4677.
The F-statistic is 13.18, with a corresponding p-value of 0.002468, suggesting that the model as a whole is statistically significant.
The residual standard error is 0.5696, representing the average deviation of the observed values from the predicted values. 

But when involving other cancer types in the linear regression analysis, the result states for both drug-genes pairs that the relationship between the drug response and CNVs of the drugs target is not significant, indicated by high p-values of the F-statistics (Graphs in Appendix).

**"Differential Treatment Efficacy in Oncological and Non-Oncological Drugs: A Fold Change Analysis"**

In our study, we performed two separate fold change analyses to compare, differentiate, and evaluate the relative effectiveness of oncology and non-oncology treatments across various cell lines, first in general way and further with special attention in gastric cancer cell lines.

The initial analysis aimed to compare the fold changes in non-oncology treatments relative to oncology treatments across all cell lines.
For context, oncology-related treatments constituted 6.4% of the total treatments in the dataset, with non-oncology treatments making up the remaining 93.6% (see Fig. in Supplementary Materials).

A visual comparison of the cancer cell lines via a bar plot (Fig. ) revealed that the majority had a fold change value less than 1.
This suggests that for these cell lines, oncology treatments were generally more effective at reducing cell growth. 

However, a few cell lines demonstrated larger fold changes, both positive and negative.
Positive fold changes indicate a greater effectiveness of non-oncology treatments, while negative fold changes point to instances where the oncology treatment resulted in an increase in cell growth, rendering the non-oncology treatment more effective by comparison, however these latter instances can be seen as outliers or special cases.

To complement the fold change analysis, we performed a statistical comparison of the average responses to oncology and non-oncology treatments across all cell lines using Welch's Two Sample t-test.
This analysis aimed to determine if the differences observed in the fold change analysis were statistically significant.
Prior to this, we first tested our data for normality using the Shapiro-Wilk test, accompanied by visual inspections using histograms and QQ plots.
We obtained a p-value of 0.5333 for oncology responses and a p-value of 0.07435 for non-oncology responses, both higher than the common alpha level of 0.05, suggesting we could not reject the null hypothesis of normality.
This confirmed, both statistically and visually (see see Fig. in Supplementary Materials), that the distribution of our data was reasonably normal, thereby supporting the validity of subsequent t-tests.

Furthermore, a one-sided t-test was conducted to specifically investigate whether the mean response to oncology treatments is significantly less (more negative) than the mean response to non-oncology treatments and therefore the efficiency.
This test yielded a significant result (p-value = 1.643e-07), suggesting that the average response to oncology treatments is significantly more negative (indicating more effective at reducing cell growth) than the average response to non-oncology treatments.
This difference is also clearly visualized in the corresponding boxplot, where the median value for oncology responses is more negative than that of the non-oncology group (Figure X).
These statistically verified facts are crucial for our subsequent analyses.
They affirm the suitability of using the oncology group as a control group for further fold-change analyses.
Moreover, the statistical significance of these observed differences reinforces our premise that certain non-oncology treatments have a considerable effect against cancer.
These treatments are potential targets for further research, contributing to our goal of drug repurposing in cancer treatment

The next fold change analysis was between non-Oncology Treatments Relative to Oncology gastric Treatments.
Here the process was the same as previously mentioned but now this time only with gastric cancer cell lines and oncology gastric treatments.
The visual representation (fig. ) highlighted that all the cell lines presented a fold change value below one, which implies that oncology gastric treatments generally demonstrate higher efficacy across the 17 gastric cell lines compared to the non-oncology treatments.
Among these, three cell lines - ACH-000678, ACH-000919, and ACH-000351 - displayed fold change values between 0.7 and 0.8.
These cell lines, being the closest to 1.0, showed comparatively higher response to non-oncology treatments, and the first two were selected for further analysis in the next section.

To ensure the validity of our findings, we confirmed the normal distribution of our data (p-value of 0.9127 and 0.08854 from the Shapiro-Wilk normality test) and validated it with visual plots (refer to Supplementary Materials).
Consequently, the two-sided t-test yielded a p-value of 0.003673, affirming a statistically significant in the observed differences.

To evaluate treatment efficiency, a one-sided t test was conducted.
With a p-value of 0.001837, we found statistically significant evidence suggesting that oncology gastric treatments are more effective than their non-oncology counterparts.

This conclusion is further emphasized by the boxplot (Figure X), which contrasts the average responses of Oncology and Non-Oncology treatments for gastric diseases.
The boxplot distinctly reveals a lower median value for Oncology Gastric Treatments, indicative of superior effectiveness.
Furthermore, a larger interquartile range, extending from the lower quartile (25th percentile) to the median, is noticeable in the Oncology group relative to the non-Oncology group.
This wider dispersion indicates an increased variability in the effectiveness of Oncology treatments, yet overall, demonstrates their superior efficiency compared to non-oncology drugs.

## **Drug Responses in High Fold Change Cell Lines: A Focus on HDAC Inhibitors**

As previously discussed, the fold change analysis between non-oncology and oncology gastric treatments led to the selection of two cell lines that were closest to 1 for further investigation.
We aimed to discover the most efficient treatments by setting a response threshold of -7, implying that only treatments with values of -7 and more negative, representing high efficiency, were selected.

Our analysis revealed, through a bar plot, that only 17 treatments demonstrated this level of efficiency on the first selected cell line, ACH-000674.
The objective was to verify if any of the highly responsive treatments corresponded with those extracted from the Principal Component Analysis (PCA) described in Section XXX. We found that 14 out of the 17 selected treatments aligned with the PC1 group of treatments from the PCA, indicating a significant overlap.

Furthermore, we sought to determine if any of the selected treatments coincided with the HDAC treatments, previously identified as potential target drugs.
Remarkably, two out of the six HDAC treatments ("BRD-K61397605-001-03-4::10::HTS002" and "BRD-K61397605-001-03-4::2.5::HTS002") were among the most efficient treatments for cell line ACH-000674, reinforcing the notion of HDAC as a potential target for drug repurposing in gastric cancer. 

A subsequent boxplot comparing the responses of these two treatments against the other efficient treatments illustrated that the median line for the treatments of interest was lower than for the rest of the treatments.
This suggested a superior efficacy for the two HDAC treatments compared to non-HDAC treatments.

The same analysis was performed on the second selected cell line, ACH-000919, and upon applying the response threshold of -7, we identified 29 treatments that showed a response lower than -7, indicating high efficiency.
Of these, an impressive 27 treatments overlapped with the PC1 group from the PCA.
Further, all six HDAC treatments were among these highly efficient treatments, once again reinforcing the potential of HDAC inhibitors as promising targets for drug repurposing in gastric cancer treatment.

# **Discussion**

## Understanding and analyzing the drug-response of existing drugs in cancer cell lines especially in gastric cancer cell lines to identify possible targets and metabolic pathways is important for drug-repurposing and to identify effective treatments for cancer. 

Through our focused analysis performing PCA and Pearson correlation, we analyzed the relationships between drugs and cell growth in gastric cancer cell lines.
By performing the analysis we selected the specific drugs, which are inhibiting the cell growth  and  gained insights into their potential targets and molecular pathways..
According to our results, the majority of the targets of the selected drugs were subtypes of the HDAC enzyme family followed by MTOR and EGFR.
Moreover, the drugs which inhibit HDAC did have a strong negative effect on cell growth of gastric cancer cell lines.
HDAC is an epigenetic regulatory enzyme and removes acetyl groups from N-terminal ends from histones leading to chromatin condensation and transcriptional suppression of genes (Rajaselvi et al, 2023).
Furthermore, we wanted to investigate the correlation between drug concentration and drug response in the context of gastric cancer, specifically focusing on HDAC inhibitors.
Our analysis revealed a notable negative correlation between drug concentration and cell growth, indicating that higher drug concentrations corresponded to a stronger influence on inhibiting cell growth.
This correlation analysis provides valuable insights into the potential efficacy of HDAC inhibitors in gastric cancer and underscores the importance of dosage considerations in developing more effective therapeutic approaches for treating gastric cancer. 

Understanding the significance of HDAC subtypes in drug response, we analyzed the mutation types of HDAC genes, resulting in mostly missense and silent mutations of HDAC subtypes.
The studies have shown that particularly the missense mutations in the HDAC class 1, 2 and 3 lead to dysregulation of enzymes which is strongly correlated with tumor progression in gastric cancer.
Han.
M et al, 2019, Rajendran et al, 2011).
In our results we found that there are mostly missense mutations in HDAC1-HDAC9, which are the subtypes in HDAC class 1,2, and 3.
That is why HDAC dysregulated enzymes are an important target for drug repurposing and HDAC inhibitors are potential anticancer treatments.
Furthermore, our analysis revealed intriguing similarities between gastric cancer and breast cancer, including shared targets (EGFR, HDAC1, and mTOR) and mechanisms of action (HDAC inhibitor).
These findings hold significant promise for development of targeted therapies with potential benefits for patients in both cancer types.
Specifically, we investigated the HDAC inhibitor, which is depicted in the histogram analysis for both cancer subtypes.
This observation aligns with existing scientific literature/ research, which highlights the potential of HDAC inhibitors to reduce cell growth through their impact on critical processes such as EMT, ER reactivation, and mitotic disruption, and also their ability to induce programmed cell death in breast cancer cells.
This shared mechanism of action highlights the potential for repurposing HDAC inhibitors as targeted therapies that could benefit patients with both gastric and breast cancer.
However, further research is needed to optimize their use.
Therefore, scientists are suggesting HDAC-based multi-target inhibitors, which have enhanced anticancer potency, ability to reduce drug resistance and potential for a further development. 
(Shian. S, et al, 2023).. 

## **"Differential Treatment Efficacy in Oncological and Non-Oncological Drugs: A Fold Change Analysis"** 

Our study provides a compelling exploration into the comparative effectiveness of oncology and non-oncology treatments across various cell lines, along with a specialized focus on gastric cancer cell lines.

The initial fold change analysis, incorporating all cell lines, substantiates the superior efficacy of oncology treatments.
Nevertheless, our findings also shed light on the potential for drug repurposing, identifying a number of instances where non-oncology treatments exhibited larger positive fold changes.
Although these instances are outliers, they carry significant implications for the potential applications of non-oncology drugs in cancer therapy.

Delving deeper into gastric cancer cell lines, our data corroborates the trend of oncology treatments demonstrating a higher efficacy.
However, a subset of these cell lines showcased a stronger response to non-oncology treatments.
This observation not only highlight the importance of understanding individual cell line responses, but also shows the potential for the implementation of personalized oncology strategies, where the deployment of non-oncology treatments might offer superior efficacy in specific contexts.

Moreover, these findings underscore the importance of individual cell line characteristics in shaping the effectiveness of different treatments.
They suggest that opportunities for improved outcomes may lie with non-oncology treatments under specific circumstances.
Therefore, further investigations are warranted to comprehend the biological mechanisms driving these differences, and to probe the potential for personalized treatment strategies predicated on cell line-specific responses.

## **Further analysis fold change - cell liness** 

Our investigation into the efficacy of non-oncology versus oncology treatments in specific gastric cancer cell lines has yielded insightful findings.
We focused on cell lines ACH-000674 and ACH-000919, setting a strict response threshold of -7 to identify the most efficient treatments.

For the ACH-000674 cell line, 17 treatments exceeded our threshold, with 14 aligning with the PCA-derived PC1 treatment group.
Importantly, two of these treatments were found to target HDAC, reinforcing its potential as a therapeutic target in gastric cancer.

A parallel analysis on the ACH-000919 cell line resulted in 29 treatments surpassing our efficiency threshold, with 27 overlapping with the 1st PCA group, and all 6 HDAC inhibitors included.
A subsequent box plot comparison confirmed their superior efficacy in comparison to other treatments.

While these findings are promising, they are based on in vitro data and focus solely on drug efficiency, without consideration of potential toxicity or adverse effects.
Thus, further validation through in vivo studies and clinical trials would be required.

In summary, our approach underscores the potential of data-driven drug repurposing, identifying promising candidates potentially overlooked in traditional oncology context, and emphasising the potential of HDAC inhibitors in the treatment of gastric cancer.

## Prediction model for the drug response of a drug in gastric cancer cells based on its target gene

To investigate the relationship between a drug and its target gene in gastric cancer cells, the drug response, gene expression, and CNVs of the drug and the respective target gene were examined.
This analysis identified several drug-gene pairs, which were then subjected to further analysis using linear regression.
By employing linear regression, prediction models were created to estimate the drug response based on the gene expression or the CNVs of the target gene.
Since the whole analysis was based on gastric cancer which represents a small group of cell lines with just approximately 17 cell lines in the data set, the limitations of the achieved results should be kept in mind.
For example the correlations between  BRD-K56957086-001-06-3 and BRD-K35329391-334-01-7  and the gene expression of their target genes did show inconsistency when increasing the drug doses, which could indicate that the high correlations for some dosages were more or less coincidence and are not proving a relationship between drug response and gene expression.
These findings are based on gene expression and CNVs in the cancer cells but no control group for this data was given, meaning no insight was given on how the gene expression or the CNVs differ generally from cells without cancer.
The limited amount of tested gastric cancer cell lines and missing information about the gene expression and the CNVs in cells without cancer  make it difficult to conclude concrete facts out of this analysis.
Though the idea of analyzing the relationship between the tested drug in drug repurposing and its original target gene could be an approach to identify or exclude certain cancer types based on gene expression or CNV ´s for the application of the drug.

The created prediction models for the drug response of BRD-K56957086-001-06-3 and BRD-K35329391-334-01-7  based on the CNVs of their target gene proved to significant at least when analyzing the given summary of the linear regression analysis of the models.
But when four more cancer types were involved in the prediction model to increase the amount of data points for the analysis the results turned out be not significant anymore.
But again since the amount of data for gastric cancer cells is so little, it was not possible to check if the discrepancy occurred because these genes are only related to the drug response in gastric cancer cells or if this proofs that there is no significant correlation.
But since all five cancer types have in common that these two drugs reduced their cell growth, it indicates that most probably analyzing the CNVs of their target gene is not a good parameter to predict the drug response. 
