---
title: "Bioinfo Project"
output: html_document
date: "2023-05-11"
---

\<\<\<\<\<\<\< Updated upstream

Load the dataframes

```{r}
load("~/Uni Heidelberg Molekulare Biotechnologie/4.Semster/Bioinfo/cellline_datasets.rda")
load("~/Uni Heidelberg Molekulare Biotechnologie/4.Semster/Bioinfo/prism_datasets.rda")
```

Create a new data frame of prism ( data cleanup and data reduction)

```{r}
View(prism)
```

#Step1 : Remove the NA values from the columns for the data reduction median value impute

```{r}
prism_cleaned <- prism
prism_cleaned[is.na(prism_cleaned)] <- 0
prism_cleaned <- prism_cleaned[complete.cases(prism_cleaned), ]
sum(is.na(prism_cleaned))
dim(prism_cleaned)
```

```{r}
View(prism_cleaned)
```

#Step 2: Visualize the data.frame for future data reduction

```{r}
class(prism_cleaned)
typeof(prism_cleaned)
prism_cleaned <- (
  sapply(prism_cleaned, as.numeric)
)
is.numeric(prism_cleaned)
```

```{r}
hist(prism_cleaned)
```

#Step 3: We want to remove the drugs, which have positive effect on the cell growth. We want to create a vector which includes more than 200 positive values as column(200 celllines have a postive growth number)

```{r}
positive_cols <- colSums(prism_cleaned > 0) > 200
num_positive_cols <- sum(positive_cols)
num_positive_cols

```

#Step 4: We want to remove selected columns. Columns are the drugs.

```{r}
#Remove the drugs which contribute to positive cell growth of cancer cell lines
prism_cleaned <- prism_cleaned[, !positive_cols]
positive_values <- colSums(prism_cleaned > 0) > 0
```

```{r}
##dimension of the data
dim(prism_cleaned)
View(prism_cleaned)
```

Data visualization of prism_cleaned data frame, which include only the columns(drugs) that influence the cell growth of the cancer cell lines negatively

```{r}
heatmap(prism_cleaned)
```

# General data visualization of selected 5 cancer cell lines 

# First create 5 vectors

```{r}
# vector of breast cancer cell lines
breast.cancer.celllines <- subset(prism.cl, lineage == "breast")
```

```{r}
View(breast.cancer.celllines)
```

```{r}
#Vector of gastric cancer cell lines
gastric.cancer.celllines <- subset(prism.cl, lineage == "gastric")
```

```{r}
View(gastric.cancer.celllines)
```

```{r}
#vector of kidney cancer cell lines
kidney.cancer.celllines <- subset(prism.cl, lineage == "kidney")
```

```{r}
View(kidney.cancer.celllines)
```

```{r}
#Vector of liver cancer cell lines:
liver.cancer.celllines <- subset(prism.cl, lineage == "liver")
```

```{r}
View(liver.cancer.celllines)
```

```{r}
#Vector of bone cancer cell lines:
bone.cancer.celllines <- subset(prism.cl, lineage == "bone")
```

```{r}
View(bone.cancer.celllines)
```

#Create a new prism data frame with selected cancer cell lines
##Step 1: Extract the IDs from prism.cl and create a new data frame of prism

```{r}
gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID
```

```{r}
View(gastric.cancer.celllines.v)
```

```{r}
breast.cancer.celllines.v <- breast.cancer.celllines $DepMap_ID
```

```{r}
View(breast.cancer.celllines.v)
```

```{r}
kidney.cancer.celllines.v <- kidney.cancer.celllines $DepMap_ID
```

```{r}
View(kidney.cancer.celllines.v)
```

```{r}
bone.cancer.celllines.v <- bone.cancer.celllines $DepMap_ID
```

```{r}
View(bone.cancer.celllines.v)
```

```{r}
liver.cancer.celllines.v <- liver.cancer.celllines $DepMap_ID

View(liver.cancer.celllines.v)
```

```{r}
cancer.celllines.v <- c(liver.cancer.celllines.v,bone.cancer.celllines.v,kidney.cancer.celllines.v,  breast.cancer.celllines.v,gastric.cancer.celllines.v)

```

```{r}
prism_celllines <- prism[rownames(prism) %in% cancer.celllines.v, ]
View (prism_celllines)

```

##Step 2: Remove the NA values

```{r}
typeof(prism_celllines)
prism_celllines_numeric_2 <- (
  sapply(prism_celllines, as.numeric)
)
is.numeric(prism_celllines_numeric_2)
###removing all columns that have only missing values
prism_celllines_matrix_filtered <- prism_celllines_numeric_2[, colSums(is.na(prism_celllines_numeric_2)) < nrow(prism_celllines_numeric_2)]
###save the names of the rows so we can restore them after that
original_row_names <- rownames(prism_celllines)
###replacing the NAs by median of the columns
prism_celllines_numeric_2[is.na(prism_celllines_numeric_2)] <-median(prism_celllines_numeric_2,na.rm =TRUE)

sum(is.na(prism_celllines_numeric_2))
original_row_names <- rownames(prism_celllines)
###restoring the row names
prism_celllines_matrix_filtered1 <- matrix(prism_celllines_numeric_2, nrow = nrow(prism_celllines), dimnames = list(rownames(prism_celllines), colnames(prism_celllines)))
original_row_names <- rownames(prism_celllines)
prism_celllines_matrix_filtered1 <- matrix(prism_celllines_numeric_2, nrow = nrow(prism_celllines), dimnames = list(rownames(prism_celllines), colnames(prism_celllines)))
typeof(prism_celllines_matrix_filtered1)
class(prism_celllines_matrix_filtered1)
sum(is.na(prism_celllines_matrix_filtered1))
```

# Data reduction: PCA

```{r}
#PCA
# Principal component scores (transformed data)
pca_result_matrix_filtrated1 <- prcomp(prism_celllines_matrix_filtered1)
pca_values <- pca_result_matrix_filtrated1$x
```


```{r}
# Variance explained by each principal component
pca_variances <- pca_result_matrix_filtrated1$sdev^2  
# Variance explained by each principal component
pca_proportions <- pca_variances / sum(pca_variances)  # Proportion of variance explained by each principal component
pca_loadings <- pca_result_matrix_filtrated1$rotation  # Loadings (coefficients) of each variable on each principal component
 plot(1:length(pca_variances), pca_variances, type = "b", xlab = "Principal Component", ylab = "Variance Explained",
     main = "Scree Plot")
plot(pca_values[, 1], pca_values[, 2], xlab = "PC1", ylab = "PC2")
```



##PCA visualization
```{r}
pca_data <- data.frame(PC1 = pca_values[, 1], PC2 = pca_values[, 2],CellLine = rownames(pca_values) )

bone.cancer.celllines.group<- "red"
kidney.cancer.celllines.group <- "blue"
liver.cancer.celllines.group <- "green"
gastric.cancer.celllines.group<- "purple"
breast.cancer.celllines.group <- "orange"

pca_data$Group <- NA
pca_data$Group[pca_data$CellLine %in% bone.cancer.celllines.v] <- "Group 1"
pca_data$Group[pca_data$CellLine %in% kidney.cancer.celllines.v] <- "Group 2"
pca_data$Group[pca_data$CellLine %in% liver.cancer.celllines.v] <- "Group 3"
pca_data$Group[pca_data$CellLine %in% gastric.cancer.celllines.v] <- "Group 4"
pca_data$Group[pca_data$CellLine %in% breast.cancer.celllines.v]<-"Group 5"
```

```{r}
install.packages(ggplot2)
library(ggplot2)
```

```{r}
ggplot(pca_data, aes(x = PC1, y = PC2, color = Group)) +
  geom_point() +
  scale_color_manual(values = c(bone.cancer.celllines.group, kidney.cancer.celllines.group, liver.cancer.celllines.group, gastric.cancer.celllines.group, breast.cancer.celllines.group)) +
  labs(x = "PC1", y = "PC2", title = "PCA Scatter Plot")
```


# Get the dataframe ready for future anaysis
```{r}
##lets create a transposed matrix:
prism_celllines_matrix_transposed<- t(prism_celllines_matrix_filtered1)
sum(is.na(prism_celllines_matrix_filtered1))##there are no missing values, I am so proud
class(prism_celllines_matrix_transposed)
###do not do the following code:
prism_celllines_matrix_transposed <- as.matrix(prism_celllines_matrix_transposed)
prism_celllines_matrix_transposed_copy <- prism_celllines_matrix_transposed
medians <- apply(prism_celllines_matrix_transposed, 1, median, na.rm = TRUE)
prism_celllines_matrix_transposed_copy[is.na(prism_celllines_matrix_transposed)] <- medians[is.na(prism_celllines_matrix_transposed)]
prism_celllines_matrix_transposed[is.na(prism_celllines_matrix_transposed)] <- medians[is.na(prism_celllines_matrix_transposed)]
for (i in 1:nrow(prism_celllines_matrix_transposed)) {
  row_values <- prism_celllines_matrix_transposed[i, ]
row_median <- median(row_values, na.rm = TRUE)
prism_celllines_matrix_transposed_copy[i, is.na(prism_celllines_matrix_transposed[i, ])] <- row_median
}
sum(is.na(prism_celllines_matrix_transposed_copy))
sum(is.na(prism_celllines_matrix_transposed))
is.numeric(prism_celllines_matrix_transposed)
```

```{r}
#PCA of transposed matrix:
pca_result_matrix_transposed <- prcomp(prism_celllines_matrix_transposed)
pca_values_t <- pca_result_matrix_transposed$x # Principal component scores (transformed data)
pca_variances_t <- pca_result_matrix_transposed$sdev^2  # Variance explained by each principal component
pca_proportions_t <- pca_variances_t / sum(pca_variances_t)  # Proportion of variance explained by each principal component
pca_loadings_t <- pca_result_matrix_transposed$rotation  # Loadings (coefficients) of each variable on each principal component
plot(1:length(pca_variances_t), pca_variances_t, type = "b", xlab = "Principal Component", ylab = "Variance Explained",main = "Scree Plot")
plot(pca_values_t[, 1], pca_values_t[, 2], xlab = "PC1", ylab = "PC2")


```

#PCA for drugs for our gastric cancer cellline:

```{r}
prism_gastric_cleaned_treatments_t<- t(prism_gastric_cleaned_treatments)
sum(is.na(prism_gastric_cleaned_treatments_t))
is.numeric(prism_gastric_cleaned_treatments_t)
pca_result_gastric_cleaned_treatments_t<- prcomp(prism_gastric_cleaned_treatments_t)
pca_values_gastric_cleaned_treatments_t <- pca_result_gastric_cleaned_treatments_t$x # Principal component scores (transformed data)
pca_variances_gastric_cleaned_treatments_t <- pca_result_gastric_cleaned_treatments_t$sdev^2  # Variance explained by each principal component
pca_proportions_gastric_cleaned_treatment_t <- pca_variances_gastric_cleaned_treatments_t / sum(pca_variances_gastric_cleaned_treatments_t)  # Proportion of variance explained by each principal component
pca_loadings_gastric_cleaned_treatments_t <- pca_result_gastric_cleaned_treatments_t$rotation  # Loadings (coefficients) of each variable on each principal component
plot(1:length(pca_variances_gastric_cleaned_treatments_t), pca_variances_gastric_cleaned_treatments_t, type = "b", xlab = "Principal Component", ylab = "Variance Explained",main = "Scree Plot")
plot(pca_values_gastric_cleaned_treatments_t[, 1], pca_values_gastric_cleaned_treatments_t[, 2], xlab = "PC1", ylab = "PC2")



```

# Pearson correlation of each PCs (drugs) for gastric cancer cell lines
## Selection of PCs
```{r}
#Select the principle components with high variance
pca_proportions_gastric_cleaned_treatment_t
#Select the first 6 Principle components since their contribution to variance>10 percent.
selected_components <- pca_values_gastric_cleaned_treatments_t[, 1:6]
#Calculate the pearson correlation
correlation_gastric_cancercelllines_selected_PCs <- cor(selected_components, prism_gastric_cleaned_treatments_t, method = "pearson")
correlation_gastric_cancercelllines_selected_PCs
summary(correlation_gastric_cancercelllines_selected_PCs)
```

### Visualization of PCs 
```{r}
#Visualize the pearson correlation of PCs with the gastric cancer cell lines (not so good, the next chunk is better)
barplot(correlation_gastric_cancercelllines_selected_PCs, main = "Pearson Correlation of Principal Components with Gastric Cancer Cell Lines", 
        xlab = "Principle Components", ylab = "Pearson Correlation", 
        col = "steelblue", ylim = c(-1, 1))
#create an intermediate object

```
###Analysis for specific cell lines
```{r}
gastric_cancer_cell_line1 <- prism_gastric_cleaned_treatments_t[, 1]

# Calculate the Pearson correlation between the first principal component and the first gastric cancer cell line
correlation_gastric_cancercellline1_selected_PCs <- cor(selected_components, gastric_cancer_cell_line1, method = "pearson")
correlation_gastric_cancercellline1_selected_PCs

```

```{r}
# Create a vector of correlation values
correlation_values <- c(-0.922615177, 0.083180285,-0.063771576 , 0.001032055, -0.133581302, 0.001307767, -0.186299116)

# Create a bar plot
barplot(correlation_values, 
        main = "Pearson Correlation of PCs with Gastric Cancer Cell Line 1",
        xlab = "Principal Component",
        ylab = "Pearson Correlation (R)",
        col = "steelblue")

# Adding labels to the bars
text(x = 1:7, y = correlation_values, 
     labels = round(correlation_values, 2), pos = 3)
```

# 

```{r}
# cell line 2

gastric_cancer_cell_line2 <- prism_gastric_cleaned_treatments_t[,2]

# Calculate the Pearson correlation between the first principal component and the 2. gastric cancer cell line
correlation_gastric_cancercellline2_selected_PCs <- cor(selected_components, gastric_cancer_cell_line2, method = "pearson")
correlation_gastric_cancercellline2_selected_PCs
# Create a vector of correlation values
correlation_values <- c(-0.950194073, -0.095496631, 0.154186354, -0.030713988, 0.001271816, -0.030989500, 0.066786533)

# Create a bar plot
barplot(correlation_values, 
        main = "Pearson Correlation of PCs with Gastric Cancer Cell Line 2",
        xlab = "Principal Component",
        ylab = "Pearson Correlation (R)",
        col = "steelblue")

# Adding labels to the bars
text(x = 1:7, y = correlation_values, 
     labels = round(correlation_values, 2), pos = 3)
```

```{r}
#cell line 3


gastric_cancer_cell_line3 <- prism_gastric_cleaned_treatments_t[,3]

# Calculate the Pearson correlation between the first principal component and the 3. gastric cancer cell line
correlation_gastric_cancercellline3_selected_PCs <- cor(selected_components, gastric_cancer_cell_line3, method = "pearson")
correlation_gastric_cancercellline3_selected_PCs
```

```{r}
# Create a vector of correlation values
correlation_values3 <- c(-0.927977386, 0.179019069, -0.007993429, 0.157945987, 0.036998181, -0.050914709, -0.046912095)


# Create a bar plot
barplot(correlation_values3, 
        main = "Pearson Correlation of PCs with Gastric Cancer Cell Line 3",
        xlab = "Principal Component",
        ylab = "Pearson Correlation (R)",
        col = "steelblue")

# Adding labels to the bars
text(x = 1:7, y = correlation_values3, 
     labels = round(correlation_values, 2), pos = 3)
```

```{r}
#cell line 4


gastric_cancer_cell_line4 <- prism_gastric_cleaned_treatments_t[,4]

# Calculate the Pearson correlation between the first principal component and the 4. gastric cancer cell line
correlation_gastric_cancercellline4_selected_PCs <- cor(selected_components, gastric_cancer_cell_line4, method = "pearson")
correlation_gastric_cancercellline4_selected_PCs
```

```{r}
correlation_values4 <- c(-0.94910213, 0.03838714, 0.11713865, 0.02180625, -0.04956825, -0.04882622, 0.01262041)
# Create a bar plot
barplot(correlation_values4, 
        main = "Pearson Correlation of PCs with Gastric Cancer Cell Line 4",
        xlab = "Principal Component",
        ylab = "Pearson Correlation (R)",
        col = "steelblue")

# Adding labels to the bars
text(x = 1:7, y = correlation_values3, 
     labels = round(correlation_values, 2), pos = 3)

```

```{r}
install.packages("ggplot2")
library(ggplot2)
```

```{r}
library(ggplot2)
```

#Visualization of Pearson correlation of each PC(correct version)
```{r}
# Create a data frame with the correlation values
correlation_data <- data.frame(CellLine = 1:17,
                               PC1 = c(-0.922615177, -0.950194073,   -0.927977386, -0.94910213, -0.9152670, -0.854801869, -0.95349076, -0.92154645, -0.92013903, -0.77537470,-0.950086951, -0.93543329, -0.93709137, -0.9250477213, -0.950361834,-0.93871668, -0.94822669),  
                               PC2 = c(0.083180285, -0.095496631,  0.179019069,   0.03838714, -0.1386250, 0.345058466, -0.11534765,  0.20474041, -0.06588813,  0.43218037, -0.088239721, -0.14684554, -0.00350409,  0.1180914557, -0.065329501,-0.02260016, -0.06579373), 
                               PC3= c(-0.063771576,  0.154186354, -0.007993429,  0.11713865, -0.2703927,-0.152188017, -0.07227995,  0.04222941,  0.04007063,  0.16919529, 0.145417339,  0.09744200, -0.06908235, -0.0006997035, -0.025131615,-0.04229883, -0.04697155),
                              PC4= c(0.001032055, -0.030713988,  0.157945987,  0.02180625,  0.1161595,-0.319057038, -0.02522252,  0.06795268, -0.12543977,  0.14860319,-0.034134148, -0.03687671,  0.10518056,  0.1658635330,  0.007893927,-0.04026534, -0.04409653),
                              PC5= c(-0.133581302,  0.001271816,  0.036998181, -0.04956825, -0.1377278, 0.001838434,  0.04777600, -0.05017946, -0.21094701, -0.05346703,0.001838434,  0.04777600, -0.05017946, -0.21094701, -0.05346703,0.20053979,  0.17713578),
                              PC6= c(0.001307767, -0.030989500, -0.050914709, -0.04882622,  0.0351149,-0.103958062,  0.04898928, -0.04865855,  0.21335588,  0.19243498,0.017044198, -0.13232765, -0.12824958,  0.0502075589, -0.159971620, 0.09320630,  0.11132276))
                              
                               
# Convert the data from wide to long format
correlation_data_long <- tidyr::gather(correlation_data, PC, Correlation, -CellLine)

# Create the bar plot using ggplot
ggplot(correlation_data_long, aes(x = PC, y = Correlation, fill = factor(CellLine))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Pearson Correlation of Principal Components with Gastric Cancer Cell Lines",
       x = "Principal Components",
       y = "Pearson Correlation") +
  scale_fill_discrete(name = "Cell Line") +
  ylim(-1, 1) +
  theme_minimal()
```

# Selection of the drugs which contribute positive to PC1
```{r}
#PC1 is the one with negative correlation for each 17 gastric cancer cell line:
##Let's find the drugs which contribute more to PC1:
###select the drugs which contribute positive to PC1
pc1_positive_drugs <- rownames(pca_values_gastric_cleaned_treatments_t)[pca_values_gastric_cleaned_treatments_t[, "PC1"] > 0]
pc1_positive_drugs


```

```{r}
View(prism.snv)
```


```{r}
mean(pca_loadings_gastric_cleaned_treatments_t[,1])
#### There are 678 drugs which contribute positive to the PC1. So lets select a threshold and find selected drugs.
pc1_selected_drugs <- rownames(pca_values_gastric_cleaned_treatments_t)[pca_values_gastric_cleaned_treatments_t[, "PC1"] > 20]
pc1_selected_drugs
```

##Create a new data frame with selected drugs

```{r}
# create a new data frame of prism.treat which include only those 66 drugs 
## Create a vector of these drugs
prism_treat_new <- prism.treat[rownames(prism.treat) %in% pc1_selected_drugs, ]
prism_treat_new
```

```{r}
View(prism_treat_new)
```

```{r}
library(ggplot2)
```

####Do not click on this chunk: There is a problem

```{r}
#visualize the mechanism of action (moa) of different drugs to specify which target is playing a major role by inhibiting the cell growth of gastric cancer cell lines

mechanism_counts <- table(prism_treat_new$moa)

# Step 2: Create a pie chart
pie_chart <- ggplot(data = data.frame(mechanism = names(mechanism_counts), count = as.numeric(mechanism_counts)),
                    aes(x = "", y = count, fill = mechanism)) +
  geom_col(width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(title = "Mechanism of Action Distribution")

# Display the pie chart
print(pie_chart)

```

```{r}
ggsave(filename = "pie_chart.png", plot = pie_chart, width = 12, height = 8, dpi = 300)

```

### Distribution of moa of selected drugs

```{r}
library(ggplot2)

# Create the bar chart with modified x-axis labels
bar_chart <- ggplot(data = prism_treat_new, aes(x = moa)) +
  geom_bar(fill = "steelblue") +
  theme_minimal() +
  labs(title = "Mechanism of Action Distribution", x = "Mechanism of Action", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(margin = margin(t = 10))) # Adjust the margin as needed

# Display the bar chart
bar_chart
```

###Distribution of targets from selected drugs
```{r}
# Do the same with the targets

bar_chart <- ggplot(data = prism_treat_new, aes(x = target)) +
  geom_bar(fill = "steelblue") +
  theme_minimal() +
  labs(title = "Target Distribution", x = "Target", y = "Count") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylim(0, 10)  # Adjust the range as needed

# Display the bar chart
bar_chart
```

```{r}
## Open the image to see the distribution better
ggsave("bar_chart.png", bar_chart, width = 20, height = 20, units = "cm")
```

```{r}
# Seperate the strings of the column target
big_string <- paste(prism_treat_new$target, collapse = ",")
```

```{r}
# Create a vector with targets
target_vector <- unlist(strsplit(big_string, ","))
target_vector <- gsub(" ", "", target_vector)
target_frequency <- as.data.frame(table(target_vector))
target_frequency
View(target_frequency)
```

```{r}
# Create a bar plot using ggplot2
bar_plot <- ggplot(target_frequency, aes(x = target_vector, fill = target_vector)) +
  geom_bar() +
  labs(x = "Target", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Target Frequencies")


print(bar_plot)
```

```{r}
# Install and load the stringr package
install.packages("stringr")
library(stringr)
```

```{r}
# Create a bar plot using ggplot2
target_frequency <- target_frequency[target_frequency$target_vector != "NA", ]
bar_plot <- ggplot(target_frequency, aes(x = reorder(target_vector, -Freq), y = Freq, fill = target_vector)) +
  geom_bar(stat = "identity") +
  labs(x = "Target", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylim(0, 15)  # Adjust the range as needed

# Adjust the x-axis labels
bar_plot <- bar_plot + scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))


print(bar_plot)
```

```{r}
target_frequency <- target_frequency[target_frequency$target_vector != "NA", ]
bar_plot <- ggplot(target_frequency, aes(x = reorder(target_vector, -Freq), y = Freq, fill = target_vector)) +
  geom_bar(stat = "identity") +
  labs(x = "Target", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(size = 5 ,angle = 45, hjust = 1)) +
  ylim(0, 15)  # Adjust the range as needed

# Adjust the x-axis labels
bar_plot <- bar_plot + scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))

# Remove legend for x-axis
bar_plot <- bar_plot + guides(fill = FALSE)

print(bar_plot)
```

```{r}
target_frequency <- target_frequency[target_frequency$target_vector != "NA", ]
bar_plot <- ggplot(target_frequency, aes(x = reorder(target_vector, -Freq), y = Freq, fill = target_vector)) +
  geom_bar(stat = "identity") +
  labs(x = "Target", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank()) +
  ylim(0, 15)  # Adjust the range as needed

print(bar_plot)


```

```{r}
## Open the image to see the distribution better
ggsave("bar_plot.png", bar_plot, width = 20, height = 20, units = "cm")
```

```{r}
# create a new dataframe of target frequency and select those which have more than 2 freq:

target_freq_dataframe_selected <- subset(target_frequency, Freq > 2)


# Print the new dataframe
print(target_freq_dataframe_selected)
```

```{r}
# Remove the row with "NA" in target_vector column
target_freq_dataframe_selected <- target_freq_dataframe_selected[target_freq_dataframe_selected$target_vector != "NA", ]

# Print the updated dataframe
print(target_freq_dataframe_selected)
```

# Bar plot of targets
```{r}
# Create a bar plot using ggplot2
bar_plot <- ggplot(target_freq_dataframe_selected, aes(x = reorder(target_vector, -Freq), y = Freq, fill = target_vector)) +
  geom_bar(stat = "identity") +
  labs(x = "Target", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylim(0, 15)  # Adjust the range as needed

# Adjust the x-axis labels
bar_plot <- bar_plot + scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))


print(bar_plot)
```

```{r}
#The same with moa
# Seperate the strings of the column target
big_string <- paste(prism_treat_new$moa, collapse = ",")

```

```{r}
moa_vector <- unlist(strsplit(big_string, ","))
moa_vector <- gsub(" ", "", moa_vector)
moa_frequency <- as.data.frame(table(moa_vector))
moa_frequency
View(moa_frequency)
```

```{r}
# create a new dataframe of target frequency and select those which have more than 1 freq:

moa_freq_dataframe_selected <- subset(moa_frequency, Freq > 1)


# Print the new dataframe
print(moa_freq_dataframe_selected)
```

#Bar plot of moa 
```{r}
# Create a bar plot of moa
bar_plot_moa <- ggplot(moa_freq_dataframe_selected, aes(x = reorder(moa_vector, -Freq), y = Freq, fill = moa_vector)) +
  geom_bar(stat = "identity") +
  labs(x = "moa", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylim(0, 15)  # Adjust the range as needed

##Adjust the x-axis labels
bar_plot_moa <- bar_plot_moa + scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))


print(bar_plot_moa)
```

```{r}
var_sd <- apply(prism_celllines_matrix_filtered1, 2, sd)
zero_sd_vars <- names(var_sd[var_sd == 0])
prism_celllines_filtered <- prism_celllines_matrix_filtered1[, -which(colnames(prism_celllines_matrix_filtered1) %in% zero_sd_vars)]
selected_pcs <- pca_values[, 1:8]
pc_correlation <- cor(selected_pcs, prism_celllines_filtered)
```
