---
title: "Bioinfo Project"
output: html_document
date: "2023-05-11"
---

```{r}

```
1. Dataframe gastric Cancer: Prism

Step1: Filter Gastric cancer cell lines from Prism.cl into new dataframe called gastric.cancer.celllines
```{r}

gastric.cancer.celllines <- subset(prism.cl, lineage == "gastric")

View(gastric.cancer.celllines)


```
Step2: Creating a vector including the DepMap ID from the gastric cancer cell lines as elements of the vector
```{r}
gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID

View(gastric.cancer.celllines.v)

```
Step3: Creating a new dataframe by extracting just the gastric cancer cell lines shown in the vector created in step 2

```{r}
prism_gastric <- prism[rownames(prism) %in% gastric.cancer.celllines.v, ]
View (prism_gastric)



```


```{r}
sum (is.na (prism_gastric)) 
```
Removing treatments with NA-values to clean the created data frame 
```{r}
prism_gastric_cleaned_treatments <- prism_gastric[, colSums(is.na(prism_gastric)) == 0]
View (prism_gastric_cleaned_treatments)
```



```{r}
#????????? why do we need scale and transposing?
PRISM_gastric_scaled <- t(scale(t(prism_gastric_cleaned_treatments)))

heatmap(PRISM_gastric_scaled)



```

```{r}
?scale
```
2. Dataframe Gastric Cancer: Prism.exp

```{r}
#Step3
prism.exp_gastric <- prism.exp[rownames(prism.exp) %in% gastric.cancer.celllines.v, ]
View (prism.exp_gastric)

```
3. Dataframe Gastric Cancer: Prism.cnv
```{r}
prism.cnv_gastric <- prism.cnv[rownames(prism.cnv) %in% gastric.cancer.celllines.v, ]
View (prism.cnv_gastric)
```
4. Dataframe Gastric Cancer: Prism.achilles
```{r}
prism.achilles_gastric <- prism.achilles[rownames(prism.achilles) %in% gastric.cancer.celllines.v, ]
View (prism.achilles_gastric)
```

```{r}

```

