---
title: "Bioinfo Project"
output: html_document
date: "2023-05-11"
---

<<<<<<< Updated upstream
```{r}
load("~/Uni Heidelberg Molekulare Biotechnologie/4.Semster/Bioinfo/cellline_datasets.rda")
load("~/Uni Heidelberg Molekulare Biotechnologie/4.Semster/Bioinfo/prism_datasets.rda")
```


```{r}
View(prism)
```

Step1 : Remove the NA values from the columns for the data reduction
median value impute 
```{r}
prism_cleaned <- prism
prism_cleaned[is.na(prism_cleaned)] <- 0
prism_cleaned <- prism_cleaned[complete.cases(prism_cleaned), ]
sum(is.na(prism_cleaned))
dim(prism_cleaned)
```

```{r}
View(prism_cleaned)
```

Step 2: Visualize the data.frame for future data reduction

```{r}
class(prism_cleaned)
typeof(prism_cleaned)
prism_cleaned <- (
  sapply(prism_cleaned, as.numeric)
)
is.numeric(prism_cleaned)
```
```{r}
hist(prism_cleaned)
```

Step 3: We want to remove the drugs, which have positive effect on the cell growth. We want to create a vector which includes more than 200 positive values as column(200 celllines have a postive growth number)

```{r}
positive_cols <- colSums(prism_cleaned > 0) > 200
num_positive_cols <- sum(positive_cols)
num_positive_cols

```
Step 4: We want to remove those columns
```{r}
prism_cleaned <- prism_cleaned[, !positive_cols]
positive_values <- colSums(prism_cleaned > 0) > 0
```

```{r}
dim(prism_cleaned)
View(prism_cleaned)
```


```{r}
heatmap(prism_cleaned)
```


Step 1: Create 5 vectors for the 5 different cancer cell lines


Vector of breast cancer cell lines: 
```{r}
breast.cancer.celllines <- subset(prism.cl, lineage == "breast")

View(breast.cancer.celllines)
```

Vector of gastric cancer cell lines:
```{r}
gastric.cancer.celllines <- subset(prism.cl, lineage == "gastric")

View(gastric.cancer.celllines)
```

Vector of kidney cancer cell lines:
```{r}
kidney.cancer.celllines <- subset(prism.cl, lineage == "kidney")

View(kidney.cancer.celllines)
```

Vector of liver cancer cell lines:
```{r}
liver.cancer.celllines <- subset(prism.cl, lineage == "liver")

View(liver.cancer.celllines)

```

Vector of bone cancer cell lines:
```{r}
bone.cancer.celllines <- subset(prism.cl, lineage == "bone")

View(bone.cancer.celllines)
```


Extract these IDs from prism
```{r}
gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID

View(gastric.cancer.celllines.v)
```

```{r}
breast.cancer.celllines.v <- breast.cancer.celllines $DepMap_ID

View(breast.cancer.celllines.v)
```

```{r}
kidney.cancer.celllines.v <- kidney.cancer.celllines $DepMap_ID

View(kidney.cancer.celllines.v)
```

```{r}
bone.cancer.celllines.v <- bone.cancer.celllines $DepMap_ID

View(bone.cancer.celllines.v)
```

```{r}
liver.cancer.celllines.v <- liver.cancer.celllines $DepMap_ID

View(liver.cancer.celllines.v)
```

```{r}
cancer.celllines.v <- c(liver.cancer.celllines.v,bone.cancer.celllines.v,kidney.cancer.celllines.v,  breast.cancer.celllines.v,gastric.cancer.celllines.v)

```


```{r}
prism_celllines <- prism[rownames(prism) %in% cancer.celllines.v, ]
View (prism_celllines)

```



Step 2 NA values:

```{r}
typeof(prism_celllines)
prism_celllines_numeric <- (
  sapply(prism_celllines, as.numeric)
)
is.numeric(prism_celllines_numeric)

prism_celllines_numeric[is.na(prism_celllines_numeric)] <-median(prism_celllines_numeric,na.rm =TRUE)

sum(is.na(prism_celllines_numeric))
View(prism_celllines_numeric)
```


PCA of this reduced data frame: 
```{r}
standardized_data.p <- scale(prism_cleaned)
pca_result.p <- prcomp(standardized_data.p, scale. = FALSE)
View(pca_result.p)
pcs <- pca_result.p$x
variance_explained <- pca_result.p$sdev^2 / sum(pca_result.p$sdev^2)

loadings <- pca_result.p$rotation
scree_plot <- plot(1:length(variance_explained), variance_explained, type = "b",
+                    xlab = "Principal Component", ylab = "Variance Explained")


```



Second Version: we want to impute the NA values with the mean values of each column : 

```{r}
install.packages("dplyr")
library(dplyr)

prism_cleaned <- prism %>%
  mutate(across(everything(), ~if_else(is.na(.), mean(., na.rm = TRUE), .)))
sum(is.na(prism_cleaned))
dim(prism_cleaned)
```

Step1: Filter Gastric cancer cell lines from Prism.cl into new dataframe called gastric.cancer.celllines
```{r}

gastric.cancer.celllines <- subset(prism.cl, lineage == "gastric")

View(gastric.cancer.celllines)


```
Step2: Creating a vector including the DepMap ID from the gastric cancer cell lines as elements of the vector
```{r}
gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID

View(gastric.cancer.celllines.v)

```
Step3: Creating a new dataframe by extracting just the gastric cancer cell lines shown in the vector created in step 2

```{r}
prism_gastric <- prism[rownames(prism) %in% gastric.cancer.celllines.v, ]
View (prism_gastric)



```


```{r}
sum (is.na (prism_gastric)) 
```
Removing treatments with NA-values to clean the created data frame 
```{r}
prism_gastric_cleaned_treatments <- prism_gastric[, colSums(is.na(prism_gastric)) == 0]
View (prism_gastric_cleaned_treatments)
```



```{r}
#????????? why do we need scale and transposing?
PRISM_gastric_scaled <- t(scale(t(prism_gastric_cleaned_treatments)))

heatmap(PRISM_gastric_scaled)



```

```{r}
?scale
```
2. Dataframe Gastric Cancer: Prism.exp

```{r}
#Step3
prism.exp_gastric <- prism.exp[rownames(prism.exp) %in% gastric.cancer.celllines.v, ]
View (prism.exp_gastric)

```
3. Dataframe Gastric Cancer: Prism.cnv
```{r}
prism.cnv_gastric <- prism.cnv[rownames(prism.cnv) %in% gastric.cancer.celllines.v, ]
View (prism.cnv_gastric)
```
4. Dataframe Gastric Cancer: Prism.achilles
```{r}
prism.achilles_gastric <- prism.achilles[rownames(prism.achilles) %in% gastric.cancer.celllines.v, ]
View (prism.achilles_gastric)
```

=======
# Reducing the gene data:

```{r}
#genes.snv_v is a vector holding all the Hugo_symbols of the damaging gene mutations found in gastric cancer celllines
## now this vector is going to be used to select these genes from all the genes in the dataframe prism.exp, prism.cnv and prism.achilles because these are probably correlated to the drug response 
genes.snv_v <- prism.snv_gastric$Hugo_Symbol

View(genes.snv_v)
r.prism.exp_g <- prism.exp_gastric[ ,colnames(prism.exp_gastric) %in% genes.snv_v]
r.prism.cnv_g <- prism.cnv_gastric[ ,colnames(prism.cnv_gastric) %in% genes.snv_v]
r.prism.achilles_g <- prism.achilles_gastric[ ,colnames(prism.achilles_gastric) %in% genes.snv_v]
View (r.prism.achilles_g)

```


```{r}

gastric.cancer.celllines <- subset(prism.cl, lineage == "gastric")

```

```{r}
breast.cancer.celllines <-subset(prism.cl, lineage == "breast")
breast.cancer.celllines.v <- breast.cancer.celllines $DepMap_ID

prism_breast <- prism[rownames(prism) %in% breast.cancer.celllines.v, ]
prism_breast_cleaned_drugs<- prism_breast[, colSums(is.na(prism_breast)) == 0]

class(prism_gastric_cleaned_drugs)
typeof(prism_gastric_cleaned_drugs)

prism_gastric_cleaned_drugs <- (
  sapply(prism_gastric_cleaned_drugs, as.numeric)
)
is.numeric(prism_gastric_cleaned_drugs)
mean.breast = mean(prism_gastric_cleaned_drugs)
```


```{r}
hist(prism_cleaned_drugs)
mean(prism_cleaned_drugs)
```
```{r}
View(gastric.cancer.celllines)
```

Step2: Creating a vector including the DepMap ID from the gastric cancer cell lines as elements of the vector

```{r}
gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID
```

```{r}
View(gastric.cancer.celllines.v)
```

Step3: Creating a new dataframe by extracting just the gastric cancer cell lines shown in the vector created in step 2
```{r}
prism_gastric <- prism[rownames(prism) %in% gastric.cancer.celllines.v, ]
View (prism_gastric)
```


```{r}
prism_gastric_cleaned_drugs<- prism_gastric[, colSums(is.na(prism_gastric)) == 0]
View (prism_gastric_cleaned_drugs)
```

```{r}
#plot a histogram of all gastric cancer cell lines
hist(prism_gastric_cleaned_drugs)
```




```{r}
mean(prism_gastric_cleaned_drugs)
median(prism_gastric_cleaned_drugs)
```

```{r}
negative_growth_cols <- which(prism_gastric_cleaned_treatments < -6.5, arr.ind = TRUE)[, "col"]
drugs_with_negative_growth <- colnames(prism_gastric_cleaned_treatments)[negative_growth_cols]
drugs_with_negative_growth
```

Create a new dat.frame with the rows as gastric cancer cell lines and the drugs with the drugs_with_negative_growth

```{r}
prism_gastric_selected_drugs <- prism_gastric_cleaned_treatments[, drugs_with_negative_growth]
View (prism_gastric_selected_drugs)
```

```{r}
install.packages("pheatmap")
library(pheatmap)

pheatmap(prism_gastric_selected_drugs,show_rownames = FALSE, show_colnames = FALSE)
```

```{r}
heatmap(prism_gastric_selected_drugs, col = color_scale, main = "Heatmap")
color_legend(max_value, min_value, color_scale)

heatmap(prism_gastric_selected_drugs)
```

```{r}

hist(prism_gastric_selected_drugs)


```


```{r}
#????????? why do we need scale and transposing?
PRISM_gastric_scaled <- t(scale(t(prism_gastric_cleaned_treatments)))

heatmap(PRISM_gastric_scaled)



```

```{r}
?scale
```

2.  Dataframe Gastric Cancer: Prism.exp

```{r}
#Step3
prism.exp_gastric <- prism.exp[rownames(prism.exp) %in% gastric.cancer.celllines.v, ]
```


```{r}
View (prism.exp_gastric)

```

```{r}
sum (is.na (prism.exp_gastric))
class(prism.exp_gastric)
typeof(prism.exp_gastric)

prism.exp_gastric <- (
  sapply(prism.exp_gastric, as.numeric)
)
is.numeric(prism.exp_gastric)
is.matrix(prism.exp_gastric)
General visualization
```{r}
breast.cancer.celllines <-subset(prism.cl, lineage == "breast")
breast.cancer.celllines.v <- breast.cancer.celllines $DepMap_ID

prism_breast <- prism[rownames(prism) %in% breast.cancer.celllines.v, ]
prism_breast_cleaned_drugs<- prism_breast[, colSums(is.na(prism_breast)) == 0]

class(prism_gastric_cleaned_drugs)
typeof(prism_gastric_cleaned_drugs)

prism_gastric_cleaned_drugs <- (
  sapply(prism_gastric_cleaned_drugs, as.numeric)
)
is.numeric(prism_gastric_cleaned_drugs)
mean.breast = mean(prism_gastric_cleaned_drugs)
```
```{r}
prism_cleaned_drugs<- prism[, colSums(is.na(prism)) == 0]
prism_cleaned_drugs <- ( sapply(prism, as.numeric)
)
is.numeric(prism_cleaned_drugs)

```
```{r}
hist(prism_cleaned_drugs)
mean(prism_cleaned_drugs)

```

```{r}
View(gastric.cancer.celllines)
```

Step2: Creating a vector including the DepMap ID from the gastric cancer cell lines as elements of the vector

```{r}
gastric.cancer.celllines.v <- gastric.cancer.celllines $DepMap_ID
```

```{r}
View(gastric.cancer.celllines.v)
```

Step3: Creating a new dataframe by extracting just the gastric cancer cell lines shown in the vector created in step 2

```{r}
prism_gastric <- prism[rownames(prism) %in% gastric.cancer.celllines.v, ]
View (prism_gastric)



```

```{r}
sum (is.na (prism_gastric)) 
```

Removing treatments with NA-values to clean the created data frame

```{r}
prism_gastric_cleaned_drugs<- prism_gastric[, colSums(is.na(prism_gastric)) == 0]
View (prism_gastric_cleaned_drugs)
```

```{r}
#make the data.frame numeric 
class(prism_gastric_cleaned_drugs)
typeof(prism_gastric_cleaned_drugs)

prism_gastric_cleaned_drugs <- (
  sapply(prism_gastric_cleaned_drugs, as.numeric)
)
is.numeric(prism_gastric_cleaned_drugs)




```

```{r}
#plot a histogram of all gastric cancer cell lines
hist(prism_gastric_cleaned_drugs)
```

Selection of the drugs(columns), which contribute to the negative cell growth \<-6.5
```{r}
mean(prism_gastric_cleaned_drugs)
median(prism_gastric_cleaned_drugs)
```

```{r}
negative_growth_cols <- which(prism_gastric_cleaned_treatments < -6.5, arr.ind = TRUE)[, "col"]
drugs_with_negative_growth <- colnames(prism_gastric_cleaned_treatments)[negative_growth_cols]
drugs_with_negative_growth
```

Create a new dat.frame with the rows as gastric cancer cell lines and the drugs with the drugs_with_negative_growth

```{r}
prism_gastric_selected_drugs <- prism_gastric_cleaned_treatments[, drugs_with_negative_growth]
View (prism_gastric_selected_drugs)
```

```{r}
install.packages("pheatmap")
library(pheatmap)

pheatmap(prism_gastric_selected_drugs,show_rownames = FALSE, show_colnames = FALSE)
```

```{r}
heatmap(prism_gastric_selected_drugs, col = color_scale, main = "Heatmap")
color_legend(max_value, min_value, color_scale)

heatmap(prism_gastric_selected_drugs)
```

```{r}

hist(prism_gastric_selected_drugs)


```


```{r}
#????????? why do we need scale and transposing?
PRISM_gastric_scaled <- t(scale(t(prism_gastric_cleaned_treatments)))

heatmap(PRISM_gastric_scaled)



```

```{r}
?scale
```

2.  Dataframe Gastric Cancer: Prism.exp

```{r}
#Step3
prism.exp_gastric <- prism.exp[rownames(prism.exp) %in% gastric.cancer.celllines.v, ]
```


```{r}
View (prism.exp_gastric)

```

```{r}
sum (is.na (prism.exp_gastric))
class(prism.exp_gastric)
typeof(prism.exp_gastric)

prism.exp_gastric <- (
  sapply(prism.exp_gastric, as.numeric)
)
is.numeric(prism.exp_gastric)
is.matrix(prism.exp_gastric)
```

```{r}
mean(prism.exp_gastric)
median(prism.exp_gastric)
hist(prism.exp_gastric)
```
```{r}
head(prism.exp_gastric)
```

```{r}
prism_exp_gastric_df <- data.frame(TSPAN6 = prism.exp_gastric)
mean_value <- mean(prism.exp_gastric$TSPAN6)
print(mean_value)

```

Create a new dat.frame with the rows as gastric cancer cell lines and the drugs with the drugs_with_negative_growth



```{r}

selected_genes_positive_cols <- which(prism.exp_gastric > 2.689977, arr.ind = TRUE)[, "col"]
selected_genes <- colnames(prism.exp_gastric)[selected_genes_positive_cols]
selected_genes
```

```{r}

prism.exp_gastric_selected_genes <-prism.exp_gastric[ ,selected_genes]

```


```{r}
install.packages("pheatmap")
library(pheatmap)
pheatmap(prism.exp_gastric_selected_genes)
```

3.  Dataframe Gastric Cancer: Prism.cnv

```{r}
prism.cnv_gastric <- prism.cnv[rownames(prism.cnv) %in% gastric.cancer.celllines.v, ]
View (prism.cnv_gastric)
```

```{r}
prism.cnv_gastric<-(sapply(prism.cnv_gastric, as.numeric))


hist(prism.cnv_gastric,breaks=20)

```


4.  Dataframe Gastric Cancer: Prism.achilles

```{r}
prism.achilles_gastric <- prism.achilles[rownames(prism.achilles) %in% gastric.cancer.celllines.v, ]
View (prism.achilles_gastric)
```

```{r}
plot(prism.achilles_gastric, main = "Density Plot")
```


>>>>>>> Stashed changes
